Challenges in Developing Great Quasi-Monte Carlo Software

### 2023-07-09 MCQMC2022ChallengesQMCSoftwareArticle_revision_1.tex
SC: obtained permission to add an image in Figure 2 and worked in some comments from Linus.

%\input FJHDef.tex
\newcommand{\AGSComment}[1]{{\color{red} #1}}
\newcommand{\FredComment}[1]{{\color{blue} Fred: #1}}
\newcommand{\SCComment}[1]{{\color{green} Sou-Cheng: #1}}
\newcommand{\YDComment}[1]{{\color{orange} Yuhan: #1}}
\newcommand{\JagsComment}[1]{{\color{violet} Jags: #1}}
\DeclareMathOperator{\disc}{disc}
\newcommand{\FJHNote}[1]{{\color{blue}Fred: #1}}

@@

\begin{figure}[t]
    \centering 
    \raisebox{0.5cm}{\includegraphics[width=0.4\textwidth]{BeamDrawing.png}}
    \includegraphics[width=0.5\textwidth]{cantileveredbeamwords.eps}
    \includegraphics[width=1\textwidth]{ldparallelbeam.eps}
    \caption{The expected deflection (upper left, image taken from~\cite{ParSee22a} with permission \SCComment{to be obtained}) of a cantilevered beam  (upper right, \SCComment{what is $n$ for this figure?}) with random Young's modulus.  QMCPy computes the answer in serial and in parallel using UM-Bridge and docker containers.  For small error tolerances, the time and sample size is $\calO(\varepsilon^{-1})$ (lower right), and parallel processing gives a speed-up factor of four to five (lower left). These plots may be reproduced using the Jupyter notebook \cite{MCQMC2022Figs}.  }
    \label{CDHJSfigbeam}
\end{figure}

Complex integrands might be best evaluated by libraries that have no direct connection to QMC software libraries.  In the uncertainty quantification (UQ) space, the recently developed UQ and Model Bridge (UM-Bridge) \cite{umbridge}, connects libraries that generate sequences of model parameters, $\bsx_1, \bsx_2, \ldots$ with libraries that evaluate models, $f(\bsx_1), f(\bsx_2), \ldots$.

We illustrate in Figure~\ref{CDHJSfigbeam} how UM-Bridge connects QMCPy with a differential equation boundary value problem solver for a cantilevered beam, with a random Young's modulus \SCComment{$m(x)$ of length $L$, assumed to be piecewise constant over three regions, each of which with a uniform distribution of uncertainty~\cite{ParSee22a}, i.e., $m(x) = \left[\begin{matrix} m_1(x), m_2(x), m_3(x) \end{matrix} \right]$ with $m_i(x) \sim \calU[1,1.2]^3$.  The beam position $x$ is discretized at $N=31$ equidistant points, and 
the vertical deflection $u(x)$ when the beam is subject to a given external load  $g(x)$ satisfies the Euler-Bernoulli partial differential equation (PDE):
$$ \frac{\partial^2}{\partial x^2}\left[ \exp(m(x)) \frac{\partial^2  u(x)}{\partial x^2}\right] = g(x),$$
%where $m(x) = \log E(x)$ is the log of an effective stiffness $E(x)$.  The  
with the following boundary conditions:
$$ \left.u(x)\right|_{x=0} = 0, \qquad 
\left.\frac{\partial u(x)}{\partial x}\right|_{x=0} = 0, \qquad
\left.\frac{\partial^2 u(x)}{\partial x^2}\right|_{x=L} = 0, \qquad  \left.\frac{\partial^3 u(x)}{\partial x^3}\right|_{x=L} = 0. $$
This PDE can be discretized with finite differences yielding a linear system $K(m) \hat{u} = \hat{g}$, where $\hat{u}$ and $\hat{g} \in \mathbb{R}^N$. }
%approximates $u(x)$ and $m(x)$ at the nodes, respectively. 
%The forward model $F:\mathbb{R}^3 \rightarrow \mathbb{R}^N $ maps $m = \left[\begin{matrix}m_1, m_2, m_3 \end{matrix}\right]^T$ to $\hat{u}$, while $Q : \mathbb{R}^N \rightarrow \mathbb{R}^2 $ selects the estimated solution at the 10th and 25th nodes, i.e., $ Q(\hat{u}) := \left[\begin{matrix} \hat{u}_{10}, \hat{u}_{25} \end{matrix} \right]^T$. The objective is to determine the distributions of $Q(F(m))$. 
QMCPy simulates the random Young's modulus and computes the expected deflection of the beam at the node points for a sequence of error tolerances using lattice rules.  By taking advantage of virtual machines and docker containers, UM-Bridge provides automatic load balancing to facilitate seamless parallel model evaluation. % the evaluation of the beam deflection can be computed in parallel in multiple cores of a single CPU to decrease execution time.

@@

\section*{Acknowledgements}\SCComment{Added this section}
The authors are grateful to the referee's constructive comments.

>>>>

%\input FJHDef.tex
\newcommand{\AGSComment}[1]{{\color{red} #1}}
\newcommand{\FredComment}[1]{{\color{blue} Fred: #1}}
\newcommand{\SCComment}[1]{{\color{green} Sou-Cheng: #1}}
\newcommand{\YDComment}[1]{{\color{orange} Yuhan: #1}}
\newcommand{\JagsComment}[1]{{\color{violet} Jags: #1}}
\newcommand{\LSComment}[1]{{\color{blue} Linus Seelinger: #1}}
\DeclareMathOperator{\disc}{disc}
\newcommand{\FJHNote}[1]{{\color{blue}Fred: #1}}

@@

\begin{figure}[t]
    \centering 
    \raisebox{0.5cm}{\includegraphics[width=0.4\textwidth]{BeamDrawing.png}}
    \includegraphics[width=0.5\textwidth]{cantileveredbeamwords.eps}
    \includegraphics[width=1\textwidth]{ldparallelbeam.eps}
    \caption{The expected deflection (upper left, image reproduced from~\cite{ParSee22a} with permission) of a cantilevered beam (upper right, \SCComment{what is $n$ for this figure?}) with random Young's modulus.  QMCPy computes the answer in serial and in parallel using UM-Bridge and docker containers.  For small error tolerances, the time and sample size is $\calO(\varepsilon^{-1})$ (lower right), and parallel processing gives a speed-up factor of four to five (lower left). These plots may be reproduced using the Jupyter notebook \cite{MCQMC2022Figs}.  }
    \label{CDHJSfigbeam}
\end{figure}

Complex integrands might be best evaluated by libraries that have no direct connection to QMC software libraries.  In the uncertainty quantification (UQ) space, the recently developed UQ and Model Bridge (UM-Bridge) \LSComment{\cite{UMBridge23}}, connects libraries that generate sequences of model parameters, $\bsx_1, \bsx_2, \ldots$ with libraries that evaluate models, $f(\bsx_1), f(\bsx_2), \ldots$   \LSComment{and optionally, their derivatives. UM-Bridge allows coupling of UQ codes to models regardless of the programming languages.}

We illustrate in Figure~\ref{CDHJSfigbeam} how UM-Bridge connects QMCPy with a differential equation boundary value problem solver for a cantilevered beam \SCComment{of length~$L$}, with a random Young's modulus \SCComment{$m(x)$, assumed to be piecewise constant over three regions, each of which with a uniform distribution of uncertainty~\cite{ParSee22a}, i.e., $m(x) = \left[\begin{matrix} m_1(x), m_2(x), m_3(x) \end{matrix} \right]$ with $m_i(x) \sim \calU[1,1.2]^3$.  The beam position $x$ is discretized at $N=31$ equidistant points, and 
the vertical deflection $u(x)$ when the beam is subject to a given external load  $g(x)$ satisfies the Euler-Bernoulli partial differential equation (PDE):
$$ \frac{\partial^2}{\partial x^2}\left[ \exp(m(x)) \frac{\partial^2  u(x)}{\partial x^2}\right] = g(x),$$
%where $m(x) = \log E(x)$ is the log of an effective stiffness $E(x)$.  The  
with the following boundary conditions:
$$ \left.u(x)\right|_{x=0} = 0, \qquad 
\left.\frac{\partial u(x)}{\partial x}\right|_{x=0} = 0, \qquad
\left.\frac{\partial^2 u(x)}{\partial x^2}\right|_{x=L} = 0, \qquad  \left.\frac{\partial^3 u(x)}{\partial x^3}\right|_{x=L} = 0. $$
This PDE can be discretized with finite differences yielding a linear system $K(m) \hat{u} = \hat{g}$, where $\hat{u}$ and $\hat{g} \in \mathbb{R}^N$. }
%approximates $u(x)$ and $m(x)$ at the nodes, respectively. 
%The forward model $F:\mathbb{R}^3 \rightarrow \mathbb{R}^N $ maps $m = \left[\begin{matrix}m_1, m_2, m_3 \end{matrix}\right]^T$ to $\hat{u}$, while $Q : \mathbb{R}^N \rightarrow \mathbb{R}^2 $ selects the estimated solution at the 10th and 25th nodes, i.e., $ Q(\hat{u}) := \left[\begin{matrix} \hat{u}_{10}, \hat{u}_{25} \end{matrix} \right]^T$. The objective is to determine the distributions of $Q(F(m))$. 
\LSComment{The beam model was written in C++ using the MUQ's model framework~\cite{MUQ}. It is one of the UM-Bridge's benchmark problem library, available as a container.}
QMCPy simulates the random Young's modulus and computes the expected deflection of the beam at the node points for a sequence of error tolerances using lattice rules.  By taking advantage of \LSComment{clusters and containers}, UM-Bridge provides automatic load balancing to facilitate seamless parallel model evaluation. \LSComment{As a result, thread-parallel QMCPy can offload costly model evaluations to remote clusters~\cite{seelinger2023lowering}.}

@@

\section*{Acknowledgements}
\SCComment{The authors are grateful to the referee's constructive comments. We also want to thank Linus Seelinger for feedback and discussion. }
